---
title: "Final_Draft"
author: "Esme, Zach, & Mandi"
date: "2/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(edld652)
library(tidyverse)
library(janitor)
library(here)
library(rio)
library(ggtext)
library(ggiraph)
library(scales)
library(gghighlight)
```

```{r load-data}
######## Load CLASS data, clean, & select columns ########################
class <- import(here("data", "Class20032019.xlsx")) %>% 
  janitor::clean_names() # CLASS Data Set

class <- class %>% 
  rename(fips = stfips,
         state_abbr = state) # make class data a fips to match grad_rates df, same with the state abbreviation.

class <- class %>% 
  select(1:3, 6, 18, 47)

######## Load grad_rates, clean, & select columns ##########################
grad_rates <- get_data("EDFacts_acgr_lea_2011_2019") %>% 
  janitor::clean_names() # graduation rates (and cohorts)

grad_rates <- grad_rates %>% 
  mutate(fips = readr::parse_number(fipst)) #rename fipst to fips

grad_rates <- grad_rates %>% 
  select(1:8, 10, 12:27, 30)

####### Merge CLASS with grad_rates df ######################################
class_grads <- left_join(grad_rates, class, by = "fips") # class df w/ grad rates&state

####### Load state_info, clean, & select columns ############################
state_info <- readr::read_csv("https://github.com/kjhealy/fips-codes/raw/master/state_fips_master.csv") #state data with fipst

state_info <- subset(state_info, select = c(1, 4, 6, 7, 9, 10))
# just want the columns with unrepeated information to class_grad that can be used later (example: regions and divisions)
state_info <- state_info %>% 
   mutate(state_name = tolower(state_name)) %>% 
    rename(state = state_name)


###### Merge class_grads with state_info data ##################################
data <- left_join(class_grads, state_info) # this has grad rates and state info
```

```{r fiscal-data}
## commenting out fiscal data other than 2011 for now ###################
## trying out visual with just one then will work to merge later ########

#fiscal10 <- get_data("NCES_CCD_fiscal_district_2010") %>% 
#  janitor::clean_names() %>% 
#  select(leaid, fipst, stabbr, totalrev, tfedrev,
#         tstrev, tlocrev, tcurinst, tcurssvc)

fiscal11 <- get_data("NCES_CCD_fiscal_district_2011") %>% 
  janitor::clean_names() %>% 
  select(leaid, fipst, stabbr, totalrev, tfedrev,
         tstrev, tlocrev, tcurinst, tcurssvc)

#fiscal12 <- get_data("NCES_CCD_fiscal_district_2012") %>% 
 # janitor::clean_names() %>% 
#  select(leaid, fipst, stabbr, totalrev, tfedrev,
 #        tstrev, tlocrev, tcurinst, tcurssvc)

#fiscal13 <- get_data("NCES_CCD_fiscal_district_2013") %>% 
 # janitor::clean_names() %>% 
  #select(leaid, fipst, stabbr, totalrev, tfedrev,
   #      tstrev, tlocrev, tcurinst, tcurssvc)

#fiscal14 <- get_data("NCES_CCD_fiscal_district_2014") %>% 
 # janitor::clean_names() %>% 
  #select(leaid, fipst, stabbr, totalrev, tfedrev,
   #      tstrev, tlocrev, tcurinst, tcurssvc)

#fiscal15 <- get_data("NCES_CCD_fiscal_district_2015") %>% 
 # janitor::clean_names() %>% 
  #select(leaid, fipst, stabbr, totalrev, tfedrev,
   #      tstrev, tlocrev, tcurinst, tcurssvc)

#fiscal16 <- get_data("NCES_CCD_fiscal_district_2016") %>% 
 # janitor::clean_names() %>% 
  #select(leaid, fipst, stabbr, totalrev, tfedrev,
   #      tstrev, tlocrev, tcurinst, tcurssvc)

#fiscal17 <- get_data("NCES_CCD_fiscal_district_2017") %>% 
 # janitor::clean_names() %>% 
  #select(leaid, fipst, stabbr, totalrev, tfedrev,
   #      tstrev, tlocrev, tcurinst, tcurssvc)

#fiscal18 <- get_data("NCES_CCD_fiscal_district_2018") %>% 
 # janitor::clean_names() %>% 
  #select(leaid, fipst, stabbr, totalrev, tfedrev,
   #      tstrev, tlocrev, tcurinst, tcurssvc)


# Merge data (state_info, grads, and fiscal11)
state_grads <- left_join(state_info, grad_rates)

# Merge state_grads with fiscal11
fiscal11 <- fiscal11 %>% 
  mutate(fips = readr::parse_number(fipst)) #rename fipst to fips

data_vizEC <- left_join(state_grads, fiscal11)

df.e <- data_vizEC %>% 
  select(1, 5:10, 12:13, 16:29, 32:38)
```

**Visual for class draft assignment** _DRAFT 1_
```{r viz-3-esme}
df.e1 <- data_vizEC %>% 
  tidyr::separate(ecd_rate, c("lower", "upper"), sep = "-") %>% 
  filter(!grepl("G|L|P", lower)) %>% 
  mutate(
    upper = ifelse(is.na(upper), lower, upper),
    lower = as.numeric(lower),
    upper = as.numeric(upper)
  ) %>% 
  rowwise() %>% 
  mutate(mean = mean(c(lower, upper))) %>% 
  ungroup()

df.e1 <- df.e1 %>% 
  group_by(state, region_name) %>% 
  mutate(support_mean = mean(tcurssvc, na.rm = TRUE))

ggplot(df.e1, aes(support_mean, mean)) +
  geom_point(color = "seagreen4") +
  gghighlight(stabbr == "OR") +
  scale_x_log10("Total Revenue for Support Services",
                labels = scales::dollar) +
  labs(y = "Averaege Graduation Rates",
       title = "Graduation Rates of Economically Disadvantaged Students",
       subtitle = "by Revenue for Support Services in Oregon",
       caption = "Data collected from 2011") +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold"),
        axis.text.y = element_text(face = "bold"),
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 11, face = "bold", hjust = 0))

### change to all_rates instead of ecd_rate
```

_DRAFT 2_
```{r try-heatmap}
df.e2 <- data_vizEC %>% 
  tidyr::separate(all_rate, c("lower", "upper"), sep = "-") %>% 
  filter(!grepl("GE|GT|LE|PS", lower)) %>% 
  mutate(
    upper = ifelse(is.na(upper), lower, upper),
    lower = as.numeric(lower),
    upper = as.numeric(upper)
  ) %>% 
  rowwise() %>% 
  mutate(mean = mean(c(lower, upper))) %>% 
  ungroup()

df.e2 %>% 
  select(6, 34, 40) %>% 
  drop_na() %>% 
  group_by(division_name) %>% 
  summarize(grads = mean(mean))


df.e2 %>% 
  group_by(division_name) %>% 
  mutate(grads = mean(mean)) %>% 
  ggplot(aes(totalrev, division_name)) +
         geom_tile(aes(fill = grads),
                   color = "seagreen4") +
  scale_x_log10("Total Revenue for Support Services",
                labels = scales::dollar) +
           coord_fixed()
```

_DRAFT 3_
```{r}
df.e3 <- data_vizEC %>% 
  tidyr::separate(all_rate, c("lower", "upper"), sep = "-") %>% 
  filter(!grepl("GE|GT|LE|PS", lower)) %>% 
  mutate(
    upper = ifelse(is.na(upper), lower, upper),
    lower = as.numeric(lower),
    upper = as.numeric(upper)
  ) %>% 
  rowwise() %>% 
  mutate(mean = mean(c(lower, upper))) %>% 
  ungroup()

df.e3 %>% 
  select(6, 34, 40) %>% 
  drop_na() %>% 
  group_by(division_name) %>% 
  summarize(totalrev = mean(mean))

df.e3 %>% 
  group_by(division_name) %>% 
  mutate(totalrev = mean(mean)) %>% 
  ggplot(aes(mean, division_name)) +
  geom_tile(aes(fill = totalrev),
            color = "cornflowerblue") +
  coord_fixed()

### First, get state averages for graduation rates
df.e3 <- df.e3 %>% 
  group_by(state) %>% 
  drop_na(mean) %>% 
  mutate(gradrate = mean(mean))

### Second, get region averages for graduation rates
df.e3 <- df.e3 %>% 
  group_by(division_name) %>% 
  mutate(gradregion = mean(gradrate))

### Third, get average revenue by state
df.e3 <- df.e3 %>% 
  group_by(state) %>% 
  drop_na(totalrev) %>% 
  mutate(statetotalrev = mean(totalrev))

## Fourth, get average revenue by region
df.e3 <- df.e3 %>% 
  group_by(division_name) %>% 
  mutate(revregion = mean(statetotalrev))

## Plot the heatmap
df.e3 %>% 
  ggplot(aes(gradregion, division_name)) +
  geom_tile(aes(fill = revregion),
            color = "cornflowerblue") +
  coord_fixed() +
      scale_fill_distiller(palette = "Greens",
                      na.value = "#transparent") +
    theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())


### make into bins the total revenue axis (to make categorical so we can plot our heatmap)
```

_DRAFT 3 TRYING NON-HEATMAP_
```{r}
df.e <- df.e %>% 
  tidyr::separate(all_rate, c("lower", "upper"), sep = "-") %>% 
  filter(!grepl("GE|GT|LE|PS", lower)) %>% 
  mutate(
    upper = ifelse(is.na(upper), lower, upper),
    lower = as.numeric(lower),
    upper = as.numeric(upper)
  ) %>% 
  rowwise() %>% 
  mutate(gradrate = mean(c(lower, upper))) %>% 
  ungroup()

# avg grads by state
df.e <- df.e %>% 
  group_by(state) %>% 
  drop_na(gradrate) %>% 
  mutate(avgstategrads = mean(gradrate))

# avg totalrev by state
df.e <- df.e %>% 
  group_by(state) %>% 
  drop_na(totalrev) %>% 
  mutate(avgtotalrev = mean(totalrev))

df.e %>%   ### look at Colors! slides for more add-ons; slide 95+
  ggplot(aes(avgtotalrev, avgstategrads)) +
  geom_point(data = filter(df.e, region_name == "Midwest"),
             color = "red") +
  geom_point(data = filter(df.e, region_name == "South"),
             color = "blue") +
  geom_point(data = filter(df.e, region_name == "West"),
             color = "green") +
  geom_point(data = filter(df.e, region_name == "Northeast"),
             color = "yellow") +
  scale_x_log10("Average Total Revenue for Support Services",
                labels = scales::dollar) +
  labs(y = "Average Graduation Rates")
```



<!-- trying something, Mandi this is the code I (Esme) showed you-->
```{r viz-code-for-part-of-Q1}
data_viz3 %>% 
  group_by(state, region_name) %>% 
  mutate(support_mean = mean(tcurssvc, na.rm = TRUE)) %>% 
  ggplot(aes(mean, state)) +
  geom_point(color = "seagreen4") +
  labs(y = "Averaege Graduation Rates",
       title = "Graduation Rates of Economically Disadvantaged Students",
       subtitle = "by state",
       caption = "Data collected from 2011") +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold"),
        axis.text.y = element_text(face = "bold"),
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 11, face = "bold", hjust = 0))
```

################################# ZACHS VIZ #############################################
For reviewer: Run each line of code in below ode chunks
```{r}
data.z <- data %>%
  separate(all_rate, c("lower", "upper"), sep = "-") %>% 
  filter(!grepl("G|L|P", lower)) %>%
  mutate(
    upper = ifelse(is.na(upper), lower, upper),
    lower = as.numeric(lower),
    upper = as.numeric(upper)
  ) %>% 
  rowwise() %>% 
  mutate(mean_rate = mean(c(lower, upper))) %>% 
  ungroup()

#data.z <- data.z %>% 
  #select(2,3,25:32, 39:41)
data.z <- data.z %>% 
  select(1:3, 25:32, 34:38)

data.z1 <- data.z %>% 
  group_by(stnam) %>% 
  summarize(state_avg = mean(mean_rate), # Collapse mean_rate by state to get only 1 mean value of grad rate for each state
            se = sd(mean_rate)/sqrt(n()))# get standard error of each state graduation average
data.z1$national_avg <- data.z1 %>% 
  summarize(national = mean(state_avg)) # getting national average for graduation rate

data.z1$se_national <- data.z1 %>% 
  summarize(se_national = sd(state_avg)/sqrt(n())) #getting SE of graduation rate for national average
```

This code chunk is the foundation of this data vizualization #1
```{r zviz-v1}
data.z1 %>% 
  mutate(stnam = fct_reorder(stnam, state_avg)) %>% 
  ggplot(aes(state_avg, stnam)) +
  geom_linerange(aes(xmin = state_avg - 1.96 * se,
                     xmax = state_avg + 1.96 * se),
                 color = "grey40") +
  geom_point() +
  theme_minimal()# shows state overall graduation rate with SE by state. . . 
```

```{r zviz-v2}
#regions_only <- data.z %>% 
  #select(3, 11, 12)
  #unique()
regions_only <- data.z %>% 
  select(4, 14, 15) %>% 
  unique()
z1.regions <- left_join(data.z1, regions_only, by = "stnam") #adding regions to df for plotting

z1.regions %>% 
  mutate(stnam = fct_reorder(stnam, state_avg)) %>% 
  ggplot(aes(state_avg, stnam, color = region_name)) +
  geom_linerange(aes(xmin = state_avg - 1.96 * se,
                     xmax = state_avg + 1.96 * se),
                 color = "grey40") +
  geom_point() +
   geom_vline(xintercept = 78.17352,
             color = "darkslateblue",
             linetype = "solid",
             size = 0.8) +
  theme_minimal() +
  labs(x = "Graduation Rate (%) - 2011",
       y = "",
       title = "State Graduation Rates",
       subtitle = "By Region",
       caption = "Vertical blue line indicates the graduation rate of the United States",
       color = "Region") 

# 2 - If I can find a way to compress class df to just one value of timehs per state (year 2011, since the grad_rate data is just 2011), I can add a shape aesthetic so that each state then has a color for region and a shape for what category of class$timehs2. Of course, we would still have the national average highlighted, would have names of states that fall below national average in once color, those that fall above in another color. 
# Want to move the legends to the top left of inside the plot itself. 
# add descriptive title, subtitle, legend name, x and y labs, caption for data source and notes about missing data.
# ideally, color code the subtitle to with region names corresponding to colors in legend
# remove the midpoint vertical lines, just leave major gridlines

```

```{r zviz-final-data}
class.z <- class %>% 
  group_by(year, state_abbr) %>% 
  summarize(avg_timehs2 = mean(timehs2) #adding in the CLASS data. The timehs2 variable relates to State Policies regarding PE time requirements for High School
)
#writexl::write_xlsx(class.z, here("data", "class.z.xlsx"))
class.z.2010 <- import(here("data", "class.z.2010.xlsx")) # data to merge into regions/graduation rates data

class.z.2010 <- left_join(z1.regions, class.z.2010) # joining
```

```{r zviz-final-plot}
class.z.2010$avg_timehs2 = as.factor(class.z.2010$avg_timehs2)
# converting variable to a factor for analysis
class.z.2010 %>% 
  mutate(stnam = fct_reorder(stnam, state_avg)) %>% 
  ggplot(aes(state_avg, stnam, color = region_name, shape = avg_timehs2)) +
  geom_linerange(aes(xmin = state_avg - 1.96 * se,
                     xmax = state_avg + 1.96 * se),
                 color = "grey40") +
  geom_point(size = 3) +
  theme_minimal() +
  geom_vline(xintercept = 78.17352,
             color = "darkslateblue",
             linetype = "solid",
             size = 0.8) +
  geom_vline(xintercept = 76.14586,
             color = "red",
             linetype = "dashed",
             size = 0.5) +
  geom_vline(xintercept = 80.17532,
             color = "red",
             linetype = "dashed",
             size = 0.5) +
  coord_cartesian(xlim = c(60, 90)) +
  geom_text(aes(x = 79, y = 42,
                label = "National Average (78.2)"),
                size = 3,
                color = "darkslateblue",
                angle = 270, vjust = 1.2) +
  geom_text(aes(x = 75.25, y = 42,
                label = "Lower Limit (76.2)"),
                size = 3,
                color = "red",
                angle = 90, vjust = 1.2) +
  geom_text(aes(x = 81, y = 42,
                label = "Upper Limit (80.2)"),
                size = 3,
                color = "red",
                angle = 270, vjust = 1.2) + 
  theme(panel.grid.major.y = element_blank(), 
        panel.grid.major.x = element_line(color = "gray85"), 
        axis.title.x = element_text(face = "bold", size = 12),
        plot.title = element_text(hjust = .0, size = 14, face = "bold"), 
        plot.subtitle = element_text(hjust = -.16), 
        plot.caption = element_text(hjust = .5, size = 10, face = "italic"),
        legend.title = element_text(size = 10, face = "bold")) +
  labs(x = "Graduation Rate (%) - 2011",
       y = "",
       title = "State Graduation Rates",
       subtitle = "By Region and High School Physical Education Policy",
       caption = "DATA NOTE, INSERT NOTE THAT SOME STATES 
       WERE OMITTED DUE TO MISSING DATA (LIST THOSE STATES",
       color = "Region",
       shape = "HS PE Policy") +
  scale_shape_discrete(labels = 
                         c("0" = "No Requirement or Recommendation",
                           "2" = "90 min/wk Minimum",
                           "3" = "90 - 150 min/wk",
                           "4" = "150 - 225 min/wk",
                           "5" = "Requires 225 min/wk")) +
  guides(color = guide_legend(order = 1),
         shape = guide_legend(order = 2))
# This plot shows graduation rate by state. Region where the state falls are colored. PE policy for that state is denoted by shape.

### Would like to find a way to make this plot include the states that are missing a value for either region or HS PE Policy that are denoted (all black if NA for region; an X if NA for HS PE Policy)

# Still need to focus on title formatting and fonts (will make standardized with other visuals at the end of project, for presentation)

```




__MY MESS AROUND PLOTS__ 
Below plots are components I am considering adding to the final plot. Have yet to figure out how to get the plot to not get congested when using the interactive functions.

Simple Interactive Plot for Testing
- This is a good, minimal example.
```{r}
gg_point = ggplot(data = class.z.2010) +
  geom_point_interactive(aes(x = state_avg, stnam, 
                             color = region_name,
                             shape = avg_timehs2,
                             tooltip = stnam, data_id = stnam))

z <- girafe(ggobj = gg_point)
if(interactive() ) print(z)
```

The following code chunk is incomplete. Ideally, this would be the final cleaned plot. Having trouble getting the interactivity to work with the aspect ratio. Would likely get rid of state labels on y-axis ... OR get rid of one of the legends and let the interactive component tell what each point corresponds to (e.g., region or HS PE time)

Would like to find a way to make this plot include the states that are missing a value for either region or HS PE Policy that are denoted (all black if NA for region; an X if NA for HS PE Policy)
```{r}
gg_point1 = class.z.2010 %>%  #THIS WORKS BUT DATA GETS PUSHED OFF AND CROWDED
  mutate(stnam = fct_reorder(stnam, state_avg)) %>% 
  ggplot(aes(state_avg, stnam, color = region_name, shape = avg_timehs2)) +
  geom_linerange(aes(xmin = state_avg - 1.96 * se,
                     xmax = state_avg + 1.96 * se),
                 color = "grey40") +
  geom_point_interactive(aes(x = state_avg, stnam, 
                             color = region_name,
                             shape = avg_timehs2,
                             tooltip = stnam, data_id = stnam)) +
  theme_minimal() +
  geom_vline(xintercept = 78.17352,
             color = "darkslateblue",
             linetype = "solid",
             size = 0.8) +
  geom_vline(xintercept = 76.14586,
             color = "red",
             linetype = "dashed",
             size = 0.5) +
  geom_vline(xintercept = 80.17532,
             color = "red",
             linetype = "dashed",
             size = 0.5) +
  coord_cartesian(xlim = c(60, 90)) +
  geom_text(aes(x = 79, y = 42,
                label = "National Average (78.2)"),
                size = 3,
                color = "darkslateblue",
                angle = 270, vjust = 1.2) +
  geom_text(aes(x = 75.5, y = 42,
                label = "Lower Limit (76.2)"),
                size = 3,
                color = "red",
                angle = 90, vjust = 1.2) +
  geom_text(aes(x = 81, y = 42,
                label = "Upper Limit (80.2)"),
                size = 3,
                color = "red",
                angle = 270, vjust = 1.2) + 
  theme(panel.grid.major.y = element_blank(), 
        panel.grid.major.x = element_line(color = "gray85"), 
        axis.title.x = element_text(face = "bold", size = 12),
        plot.title = element_text(hjust = .0, size = 14, face = "bold"), 
        plot.subtitle = element_text(hjust = -.16), 
        plot.caption = element_text(hjust = .5, size = 10, face = "italic"),
        legend.title = element_text(size = 10, face = "bold")) +
  labs(x = "Graduation Rate (%) - 2011",
       y = "",
       title = "State Graduation Rates",
       subtitle = "By Region and High School Physical Education Policy",
       caption = "DATA NOTE, INSERT NOTE THAT SOME STATES 
       WERE OMITTED DUE TO MISSING DATA (LIST THOSE STATES",
       color = "Region",
       shape = "HS PE Policy") +
  scale_shape_discrete(labels = 
                         c("0" = "No Requirement or Recommendation",
                           "2" = "90 min/wk Minimum",
                           "3" = "90 - 150 min/wk",
                           "4" = "150 - 225 min/wk",
                           "5" = "Requires 225 min/wk")) +
  guides(color = guide_legend(order = 1),
         shape = guide_legend(order = 2))

z.1 <- girafe(ggobj = gg_point1)
if(interactive() ) print(z.1) # How can I make this work. The animation and all components are included, but the plot is smashed together and crowded. 
```


####################### End of Zach Viz #####################

############# Mandi Vizualization #####################





############## End Mandi Viz ###########################